---
title: "CodeAlong"
author: "Jeanne McClure"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# CODE- ALONG #######################

## 1. Programming Basics ###########


#####################
### a. Functions ####
#####################


mean()
max()
filter()
class()



```{r}
round(3.14)
#' In the space below, use the round() function to 
#' round 3.14 to the nearest whole number.
#' Run your code to see the result. 

```


######################
### b. Arguments #####
######################

#### args() function ###############

```{r}
?mean()
?class()

```

Take a look at the arguments for the round() function. 

```{r}
?round()
```


In the code chunck below:

1. Add an argument to the round() function to 
2. round pi to the nearest 1 decimal places. 
3. Remember to add a comma between arguments.


```{r}

```


#######################
### c. Objects ########
#######################

```{r}
twenty <-20
twenty
thirty <- 20
thirty
forty <- 20
forty
new_data <- 20
new_data


```

Try to run the following in a code chunck...
13 <- 20 *4
```{r}

```

What happens?

Now run the following in a code chunk...
x <- 20 *4
```{r}

```
What happens?

Don't ever have names like TRUE/FALSE or NaN, inf, NA

####  <- Operator ##################

In the chunck below

1. Assign a number between 1 and 10 to a new object call my_number.
2. Multiply my_number by 2 and save as my_product. 
3. Type my_product and run to examine the contents. 

```{r}


```

########Other operators 

### Pipe operator

percent greater than percent

 %>%

The original pipe operator, %>%, comes from the {magrittr} package but all packages in the tidyverse load %>% for you automatically, so you don’t usually load magrittr explicitly.



 |>

Some labs will also use a very powerful |> operator called a pipe. Pipes are a powerful tool for combining a sequence of functions or processes. 

The pipe has become such a useful and much used operator in R that it is now baked into R using the new and simpler version of the pipe |> operator demonstrated in the following code chunk:

#####################
### d. Packages #####
#####################

#### install.packages() function ######
 You should do this in the console...no worries we will show you how :)


install.packages("tidyverse")


```{r}
library(tidyverse)
?tidyverse

```

Run the following code to 
install the readr package. 



did it work? Maybe you need to install first. You can also check if you package is loaded in the package pane.

Now use the library() function to load the readr package into R.
```{r}
library(readr)

```

Use ? to check the help document for the readr package.

```{r}

```

